\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{siunitx}

\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue}

\title{Hegemon Prover Architecture:\\
Formal Constraints, Security Accounting, Operation Budgets, and Feasibility}
\author{Hegemon Engineering}
\date{2026-02-28}

\begin{document}
\maketitle

\begin{abstract}
This document reformulates Hegemon proving as a constrained cryptographic systems problem. The goal is not to "optimize code" in the abstract; it is to satisfy explicit inequalities for security, latency, queue stability, and hardware budget. We derive operation counts from the current implementation constants (Poseidon2 rounds, trace dimensions, FRI settings, recursion payload shape), compute cycle budgets on \texttt{hegemon-prover} (32 cores, 256 GB RAM), and prove where the current architecture is infeasible. We then derive a replacement architecture from those constraints.
\end{abstract}

\section{Objective and Hard Constraints}

\subsection{Objective}
Sustain target throughput $T_{\mathrm{target}}$ with full post-quantum privacy, no proof bypass, and deterministic consensus import.

\subsection{Non-negotiable constraints}
\begin{enumerate}[label=C\arabic*.,leftmargin=2em]
  \item \textbf{Full PQ security:} no reduced parameters, no non-cryptographic fast path.
  \item \textbf{Fail-closed validity:} invalid proof data must be rejected at import.
  \item \textbf{Bounded pending latency:} no indefinite queue growth in steady-state.
  \item \textbf{Hardware reality:} heavy compute must run on \texttt{hegemon-prover}, not the laptop.
  \item \textbf{Acceptance gate:} sustained $\ge 10$ TPS for 30 minutes in full-PQ mode.
\end{enumerate}

\section{Implementation Constants (From Code)}

This section uses concrete constants currently in the codebase:

\begin{itemize}[leftmargin=2em]
  \item Transaction AIR rows: $N = 8192 = 2^{13}$.
  \item Trace width: $W=121$ columns.
  \item Poseidon2 parameters: width $t=12$, external rounds $R_F=8$ (split 4/4), internal rounds $R_P=22$, total steps per permutation $=31$.
  \item Merkle depth in transaction circuit: $32$.
  \item Per-transaction cycle schedule: cycle length $32$, total cycles $256$ (padded), used cycles $143$.
  \item Production FRI profile: log blowup $=4$ (blowup $b=16$), queries $q=32$.
  \item AIR max constraint degree used by quotient chunking logic: $8$, giving log quotient chunks $=\lceil\log_2(8-1)\rceil=3$, i.e. $8$ chunks.
\end{itemize}

Derived domain size for production FRI:
\begin{equation}
M = bN = 16\cdot 8192 = 131072 = 2^{17}.
\end{equation}

\section{Cryptographic Security Accounting}

\subsection{Inner proof soundness decomposition}
Write inner proof failure probability as
\begin{equation}
\varepsilon_{\mathrm{inner}} \le \varepsilon_{\mathrm{AIR}} + \varepsilon_{\mathrm{FRI}} + \varepsilon_{\mathrm{FS}} + \varepsilon_{\mathrm{hash}}.
\end{equation}
For FRI-like testing on quotient chunks with blowup $b$ and $q$ queries, the standard term is of the form
\begin{equation}
\varepsilon_{\mathrm{FRI}} \approx \left(\frac{1}{b}\right)^q
\end{equation}
up to low-order protocol constants.

For production parameters $b=16$, $q=32$:
\begin{equation}
\left(\frac{1}{16}\right)^{32}=2^{-128}.
\end{equation}
So the FRI term is in the 128-bit range before adding transcript/hash terms.

\subsection{Block-level recursion composition}
If one block aggregates $n$ statements and verifies one recursive root proof:
\begin{equation}
\varepsilon_{\mathrm{block}} \le \varepsilon_{\mathrm{outer}} + n\,\varepsilon_{\mathrm{inner}} + \varepsilon_{\mathrm{binding}} + \varepsilon_{\mathrm{hash}}.
\end{equation}
For $n\le 256$ and $\varepsilon_{\mathrm{inner}}\approx 2^{-128}$, the union term is $\le 2^{-120}$.

\subsection{Binding hash and commitment security}
Hegemon binds statements through 48-byte commitments and Blake-domain-separated hashes. With 384-bit outputs, collision resistance in a quantum model is still in the practical $\approx 2^{128}$ class (birthday-style quantum collision bounds), which is aligned with the target soundness class.

\section{Operator Cost Functions (Formal)}

Let TPS be $T$, block interval $\tau_b$, and transactions per block $n=T\tau_b$.

\subsection{Wallet}
\begin{align}
C_{\mathrm{wallet,send}} &= C_{\mathrm{select}} + C_{\mathrm{witness}} + C_{\mathrm{pack}} + C_{\mathrm{submit}},\\
C_{\mathrm{wallet,scan/sec}} &\approx T\cdot O\cdot C_{\mathrm{decrypt}},
\end{align}
where $O$ is outputs/tx.

\subsection{Coordinator}
\begin{equation}
C_{\mathrm{coord}}(m)\approx O(m\log m)+O(m\,C_{\mathrm{conflict}})+O(m\,C_{\mathrm{locality}}),
\end{equation}
for mempool size $m$.

\subsection{Prover (dominant)}
\begin{equation}
C_{\mathrm{prover}}(n)=C_{\mathrm{pre}}(n)+C_{\mathrm{inner}}(n)+C_{\mathrm{agg}}(n).
\end{equation}
Current implementation path for aggregation is monolithic in $n$ (one outer circuit that verifies all $n$ inner proofs), not a stage-tree recursion.

\subsection{Queue stability}
\begin{equation}
\frac{dQ}{dt}=\lambda-\mu,
\end{equation}
with stable operation requiring $\mu\ge\lambda$ over sustained windows.

\section{Exact Poseidon2 Operation Accounting (From Code)}

\subsection{Per-round operation counts}
From \texttt{poseidon2.rs}:
\begin{itemize}[leftmargin=2em]
  \item S-box $x\mapsto x^7$: 4 multiplications.
  \item External round (width 12):
  \begin{align}
  M_{\mathrm{ext}} &= 12\cdot 4 = 48,\\
  A_{\mathrm{ext}} &= 12\;\text{(add constants)} + 57\;\text{(mds\_light)} = 69.
  \end{align}
  \item Internal round:
  \begin{align}
  M_{\mathrm{int}} &= 4\;\text{(sbox on lane 0)} + 12\;\text{(diag mult)} = 16,\\
  A_{\mathrm{int}} &= 1\;\text{(lane-0 constant)} + 12\;\text{(sum)} + 12\;\text{(add sum)} = 25.
  \end{align}
  \item Initial linear step (step 0): $M_{0}=0$, $A_{0}=57$.
\end{itemize}

With 8 external rounds and 22 internal rounds:
\begin{align}
M_{\mathrm{perm}} &= 8\cdot 48 + 22\cdot 16 = 736,\\
A_{\mathrm{perm}} &= 57 + 8\cdot 69 + 22\cdot 25 = 1159.
\end{align}
Total field ops per permutation:
\begin{equation}
\mathrm{Ops}_{\mathrm{perm}} = M_{\mathrm{perm}} + A_{\mathrm{perm}} = 1895.
\end{equation}

\subsection{Per-transaction hash-step operations in witness generation}
The transaction trace loop runs 256 cycles, each with one Poseidon2 permutation schedule.
\begin{align}
M_{\mathrm{tx,witness}} &= 256\cdot 736 = 188{,}416,\\
A_{\mathrm{tx,witness}} &= 256\cdot 1159 = 296{,}704,\\
\mathrm{Ops}_{\mathrm{tx,witness}} &= 485{,}120.
\end{align}

Important: this is only witness-side hash algebra, \emph{not} full STARK proving.

\section{STARK Prover Lower-Bound Accounting}

\subsection{FFT/LDE lower bound}
For production domain $M=2^{17}$ and trace width $W=121$:
\begin{equation}
B_{\mathrm{trace}} = W\,M\,\log_2 M = 121\cdot 131072\cdot 17 = 269{,}615{,}104
\end{equation}
butterfly units.

Using a conservative 3 field ops per butterfly:
\begin{equation}
\mathrm{Ops}_{\mathrm{trace,LDE}}\gtrsim 808{,}845{,}312.
\end{equation}

Quotient chunk columns are 8 (from log chunks = 3), adding
\begin{equation}
\mathrm{Ops}_{\mathrm{quot,LDE}}\gtrsim 53{,}477{,}376.
\end{equation}

This is a lower bound for one transform pass; practical pipelines do multiple passes.

\subsection{Merkle commitment lower bound}
At least two major oracle trees (trace + quotient) plus FRI commit-phase trees. With $M=131072$ and observed commit-phase length $c=13$ (from recursion spike output):
\begin{align}
N_{\mathrm{nodes}}^{\mathrm{trace+quot}} &\approx 2(M-1)=262{,}142,\\
N_{\mathrm{nodes}}^{\mathrm{fri-phase}} &\approx \sum_{i=1}^{13}(M/2^i-1)=131{,}043,\\
N_{\mathrm{nodes,total}} &\approx 393{,}185.
\end{align}

If each compression uses one Poseidon2-like permutation cost envelope:
\begin{equation}
\mathrm{Ops}_{\mathrm{Merkle}}\gtrsim 393{,}185\cdot 1895=745{,}085{,}575.
\end{equation}

\subsection{Combined prover arithmetic floor}
\begin{equation}
\mathrm{Ops}_{\mathrm{prove,lb}} \gtrsim 8.09\times 10^8 + 5.35\times 10^7 + 7.45\times 10^8
\approx 1.61\times 10^9
\end{equation}
field-level operations before many runtime constants (memory movement, orchestration, serialization, recursion circuit plumbing).

\section{Cycle Budget on \texttt{hegemon-prover}}

For 32 cores at 2.5--3.0 GHz effective:
\begin{equation}
\mathrm{cycles/sec} \in [8.0,9.6]\times 10^{10}.
\end{equation}

Budget per tx:
\begin{align}
\mathrm{budget}_{1\mathrm{TPS}} &\in [8.0,9.6]\times 10^{10}\;\mathrm{cycles/tx},\\
\mathrm{budget}_{10\mathrm{TPS}} &\in [8.0,9.6]\times 10^{9}\;\mathrm{cycles/tx}.
\end{align}

Measured hot aggregation singleton is $S\approx 5.15$ s/tx, giving:
\begin{equation}
\mathrm{cycles}_{\mathrm{hot}}\in [4.12,4.94]\times 10^{11}\;\mathrm{cycles/tx}.
\end{equation}
That is $\approx 5.15\times$ over 1 TPS budget and $\approx 51.5\times$ over 10 TPS budget.

Measured cold miss equivalent $\sim 465$ s/tx is $\sim 465\times$ over 1 TPS budget.

\section{Where the Orders of Magnitude Are Hiding}

\subsection{Fact 1: not in core note hash math}
Witness-level hash ops per tx are $\sim 4.85\times 10^5$, tiny relative to total observed cycles.

\subsection{Fact 2: in proving-system and architecture constants}
Dominant terms are:
\begin{enumerate}[leftmargin=2em]
  \item LDE/FFT + Merkle commitment memory traffic and nontrivial constants.
  \item Recursion circuit setup/build and common-data preprocessing.
  \item Shape churn and cache misses.
  \item Monolithic aggregation path keyed by exact \texttt{tx\_count}.
\end{enumerate}

\subsection{Current shape-key entropy problem}
Current cache key includes
\begin{equation}
(\texttt{tx\_count},\texttt{pub\_inputs\_len},\texttt{log\_blowup},\texttt{proof\_shape}),
\end{equation}
so variable block occupancy induces many distinct expensive keys.

With observed cold build events in the hundreds of seconds, even rare misses destroy queue stability.

\section{Architectural Diagnosis (Mathematical)}

\subsection{Critical observation from code path}
Current aggregation for $n>1$ constructs one outer recursion circuit with a loop over all $n$ inner proofs. Tree metadata (arity/levels) is carried in payload, but proving work is monolithic per batch, not levelized DAG recursion.

Hence wall-time behavior is closer to:
\begin{equation}
T_{\mathrm{agg,current}}(n)\approx T_{\mathrm{cache}}(n)+T_{\mathrm{pack}}(n)+T_{\mathrm{runner}}(n)+T_{\mathrm{outer\_prove}}(n),
\end{equation}
with $T_{\mathrm{cache}}(n)$ potentially catastrophic on misses.

\subsection{Why this jams even with large hardware}
If cold events leak into critical path, effective service rate drops below arrival rate:
\begin{equation}
\mu_{\mathrm{effective}} < \lambda \Rightarrow Q(t)\ \text{diverges}.
\end{equation}
This is a control-stability failure, not just "insufficient cores".

\section{Replacement Architecture Derived from Constraints}

\subsection{Design theorem}
To make recursion scale, the system must satisfy:
\begin{enumerate}[leftmargin=2em]
  \item \textbf{Shape stationarity:} bounded shape set independent of instantaneous $n$.
  \item \textbf{Critical-path isolation:} cold preprocessing excluded from liveness path.
  \item \textbf{Levelized parallelism:} recursion executed as stage DAG (leaf/merge/root), not monolithic batch.
\end{enumerate}

\subsection{Tree-recursive model}
Let leaf capacity be $m$ tx/proof and merge arity $k$.
\begin{align}
L_0 &= \left\lceil\frac{n}{m}\right\rceil,\\
L_i &= \left\lceil\frac{L_{i-1}}{k}\right\rceil,\\
D &= \min\{d\,|\,L_d=1\}=\left\lceil\log_k L_0\right\rceil.
\end{align}

Parallel wall-time model:
\begin{equation}
T_{\mathrm{block}}\approx T_{\mathrm{pre}} + \max_j T_{\mathrm{leaf},j} + \sum_{i=1}^{D} T_{\mathrm{merge},i} + T_{\mathrm{root}}.
\end{equation}

Total work remains roughly linear in $n$, but critical path becomes logarithmic in tree depth rather than linear in proof count.

\subsection{Shape-count reduction}
Monolithic exact-$n$ caching requires up to $O(n_{\max})$ heavy shapes.
Tree design with fixed $(m,k)$ uses $O(\log_k n_{\max})$ merge shapes plus one leaf shape, dramatically reducing cold-synthesis risk.

\section{Feasibility Inequalities and Go/No-Go}

\subsection{Block-time feasibility}
For target TPS $T$ and block interval $\tau_b=6$:
\begin{equation}
T_{\mathrm{critical/block}} \le 6\ \mathrm{s},\qquad n=6T.
\end{equation}

\subsection{1 TPS gate}
Require all simultaneously:
\begin{align}
S_{\mathrm{hot}} &\le 1.0\ \mathrm{s/tx},\\
p95(S_{\mathrm{tx}}) &\le 2.0\ \mathrm{s/tx},\\
p_{\mathrm{cold,critical}} &\approx 0.
\end{align}

\subsection{10 TPS gate}
Require all simultaneously:
\begin{align}
T_{\mathrm{critical/block}} &\le 6\ \mathrm{s}\ \text{for}\ n\approx 60,\\
\mu_{30\mathrm{min}} &\ge 10\ \mathrm{TPS},\\
\sup_t\,Q(t) &\text{ bounded with finite drain time under burst tests.}
\end{align}

\subsection{Current decision}
\textbf{No-go} under present measured behavior.

Reason: measured hot and cold service times violate feasibility inequalities by large factors.

\section{Literature Review (Inspiration With Transfer Rules)}

\subsection{Transparent/PQ proof foundations}
\begin{itemize}[leftmargin=2em]
  \item STARK foundational framework \cite{stark2018}.
  \item DEEP-FRI soundness improvements \cite{deepfri2019}.
  \item Fractal transparent recursion \cite{fractal2019}.
\end{itemize}

\subsection{Modern recursive/folding line}
\begin{itemize}[leftmargin=2em]
  \item HyperNova folding architecture \cite{hypernova2023}.
  \item Neo/SuperNeo direction (small-field, pay-per-bit, PQ folding) from available metadata \cite{neo2025,superneo2026}.
\end{itemize}

\subsection{Permutation and implementation constants}
\begin{itemize}[leftmargin=2em]
  \item Poseidon2 design and round/linear-layer optimization \cite{poseidon22023}.
  \item Plonky3 recursion implementation model \cite{plonky3recbook}.
  \item RISC Zero recursion program model \cite{risc0repo,risc0docs}.
\end{itemize}

\subsection{Transfer rules for Hegemon}
From this literature, the transferable rules are:
\begin{enumerate}[leftmargin=2em]
  \item amortize expensive setup aggressively;
  \item keep recursion shapes fixed and cacheable;
  \item parallelize by proof DAG levels;
  \item treat memory bandwidth and cache locality as first-order constraints;
  \item enforce explicit soundness budgets under composition, not informal claims.
\end{enumerate}

\section{Actionable, Math-Driven Implementation Targets}

\subsection{Target envelope for first success (1 TPS)}
For $\tau_b=6$ s and $n=6$:
\begin{align}
T_{\mathrm{decode}} &\le 0.4\ \mathrm{s},\\
T_{\mathrm{commit}} &\le 1.8\ \mathrm{s},\\
T_{\mathrm{rec,critical}} &\le 2.8\ \mathrm{s},\\
T_{\mathrm{import}} &\le 0.7\ \mathrm{s},\\
T_{\mathrm{slack}} &\le 0.3\ \mathrm{s}.
\end{align}

\subsection{Architecture requirements implied by the envelope}
\begin{enumerate}[leftmargin=2em]
  \item Remove cold recursion setup from liveness path entirely.
  \item Replace exact-$n$ monolith with fixed-shape tree DAG.
  \item Promote prover cache from thread-local ownership to process-global manager with persistent hydration.
  \item Prebuild required shapes before entering throughput mode.
  \item Reject incompatible peers early (genesis/protocol/aggregation format).
\end{enumerate}

\section{Claude's Analysis: From First Principles to World-Scale Proving}

\subsection{Methodology and code corrections}

This section is the result of reading the full circuit codebase---1535 lines of
\texttt{aggregation/src/lib.rs}, the batch circuit (\texttt{circuits/batch/}),
the block commitment circuit (\texttt{circuits/block/}), the transaction AIR
(\texttt{transaction-core/src/p3\_air.rs}), the Plonky3 FRI configuration
(\texttt{p3\_config.rs}), and the dimension calculations
(\texttt{dimensions.rs}). Every number below is derived from code constants or
physics (clock frequency, memory bandwidth, silicon area), not from intuition
or analogy.

\medskip\noindent\textbf{Correction to Section~2.} The document states trace
width $W=121$. The actual codebase has \texttt{BASE\_TRACE\_WIDTH}~$=99$
columns (\texttt{COL\_STABLECOIN\_SLOT\_SEL3}$=98$, plus one) plus
\texttt{PREPROCESSED\_WIDTH}~$=33$ schedule columns
(\texttt{PREP\_CM1\_ROW}$=32$, plus one), giving $W=132$. All calculations
below use $W=132$.

\subsection{What the current architecture gets right}

Before diagnosing failures, it is important to acknowledge the design decisions
that are sound and should be preserved in any replacement.

\subsubsection{Poseidon2 over Goldilocks}

The choice of Poseidon2 (width 12, rate 6, $x^7$ S-box, 8 external $+$ 22
internal rounds) over the Goldilocks field ($p = 2^{64} - 2^{32} + 1$) is
correct for three reasons:
\begin{enumerate}[leftmargin=2em]
  \item \textbf{Algebraic efficiency:} Goldilocks admits a fast modular
        reduction ($a \cdot 2^{32} \equiv a - a \bmod 2^{32}$), giving
        $\sim$2.5 cycles per multiplication with AVX2 packing. This is
        $2$--$3\times$ faster than BN254 scalar arithmetic and $5\times$
        faster than general 256-bit fields.
  \item \textbf{PQ security:} The permutation's security reduces to algebraic
        attacks on a high-degree S-box ($d=7$) with full-width diffusion.
        No lattice or discrete-log structure is exposed.
  \item \textbf{STARK compatibility:} Goldilocks has a $2^{32}$-smooth
        multiplicative subgroup, enabling efficient NTT up to $2^{32}$-point
        transforms without field extensions.
\end{enumerate}

\subsubsection{The transaction AIR design}

The cycle schedule (32 rows/cycle, 143 used cycles, 256 total cycles padded to
$N=8192$) is well-designed:
\begin{itemize}[leftmargin=2em]
  \item Each Poseidon2 permutation needs 31 steps (1 init $+$ 8 external $+$
        22 internal), fitting cleanly in a 32-row cycle with one absorb/capture
        row.
  \item The constraint degree (max 8 from S-box interactions with selector
        flags) matches the quotient chunking parameter, yielding 8 quotient
        chunks---no wasted overhead.
  \item Preprocessed round constants (33 columns of schedule data) are
        generated once at compile time and reused across all proofs, avoiding
        any dynamic setup.
\end{itemize}

\subsubsection{The batch circuit}

The \texttt{BatchTransactionProverP3} in \texttt{circuits/batch/} is the most
important piece of correct engineering in the codebase. It proves up to 16
transactions in a single STARK proof by concatenating traces:
\begin{equation}
N_{\mathrm{batch}}(k) = (k \cdot 8192).\texttt{next\_power\_of\_two}(),
\end{equation}
giving domain sizes $\{2^{13}, 2^{14}, 2^{15}, 2^{16}, 2^{17}\}$ for
$k \in \{1, 2, 3\text{--}4, 5\text{--}8, 9\text{--}16\}$.
The AIR is identical to the single-transaction AIR applied at each slot offset.
Preprocessed traces are deterministic from $k$ and can be computed once at
startup. No circuit builder, no cache, no dynamic shapes.

\subsubsection{The block commitment circuit}

The \texttt{CommitmentBlockProverP3} (60 columns, Poseidon2 sponge over
statement hashes $+$ nullifier permutation argument) correctly binds the
block's transaction set to a single commitment. Its trace is small:
$\texttt{trace\_len}(n) = ((n+1).\texttt{next\_power\_of\_two}()) \times 32$.
For $n=60$: $2048 = 2^{11}$ rows. This circuit is fast ($<0.5$~s on any
reasonable hardware) and sound.

\subsubsection{FRI parameter selection}

Production parameters ($\log_2 b = 4$, $q = 32$) give $\varepsilon_{\mathrm{FRI}}
= (1/16)^{32} = 2^{-128}$, which is the correct 128-bit security target.
The debug/fast profile ($\log_2 b = 3$, $q = 8$) gives $2^{-24}$, appropriate
for testing. The feature-flag gating
(\texttt{stark-fast}/\texttt{plonky3-e2e}) is clean.

\subsection{What the current architecture gets wrong}

\subsubsection{Flaw 1: Recursion is the wrong tool for this throughput regime}

The aggregation circuit builds a Plonky3 recursion circuit that re-verifies $n$
inner STARK proofs inside an outer STARK. Each inner proof contains
approximately:
\begin{align}
W_{\mathrm{inner}} &\approx \underbrace{2 \times 6}_{\text{commitments}} +
  \underbrace{132 \times 2}_{\text{opened trace}} +
  \underbrace{8 \times 2}_{\text{quotient}} +
  \underbrace{13 \times 6}_{\text{FRI commits}} \nonumber\\
  &\quad+ \underbrace{32 \times (3 \times 132 + 3 \times 17 \times 6)}_{\text{query proofs}} +
  \underbrace{32 \times 13 \times (2 + 17 \times 6)}_{\text{FRI openings}} \nonumber\\
  &\approx 63{,}000 \text{ field elements per inner proof.}
\end{align}

Every one of these 63{,}000 elements becomes a symbolic witness target in
the outer circuit. The inner FRI verifier---32 queries, each with 17 Merkle
compression calls (each 31 Poseidon2 steps)---must be re-expressed as
algebraic constraints in the outer AIR. The S-box $x^7$ in each Poseidon2
round becomes a degree-7 symbolic expression, and for 32 queries:
\begin{equation}
\text{S-box evaluations} = 32 \times 17 \times (8 \times 12 + 22 \times 1)
= 32 \times 17 \times 118 = 64{,}192.
\end{equation}

The core absurdity: native STARK verification of a single transaction proof
costs $<10$~ms on one core. Re-verification inside a recursion circuit costs
seconds because every field operation becomes a constraint row. The recursion
approach pays $>50\times$ overhead.

\subsubsection{Flaw 2: The cache is thread-local with $O(n_{\max})$ key space}

The cache key \texttt{(tx\_count, pub\_inputs\_len, log\_blowup, ProofShape)}
makes the cache miss rate proportional to block occupancy variance. The
\texttt{thread\_local!} storage means cache entries are not shared across
Rayon worker threads. Cold build costs 100--465 seconds per miss.

\subsubsection{Flaw 3: The batch circuit is never used}

The complete, tested, working \texttt{BatchTransactionProverP3} is never
invoked by the aggregation pipeline. The hot path runs individual transaction
proofs through the recursion circuit instead.

\subsubsection{Flaw 4: The recursion circuit's outer trace is enormous}

For $n$ inner proofs, the outer circuit produces a trace with
$\sim n \times 100{,}000$ rows. For $n = 6$:
\begin{equation}
N_{\mathrm{outer}} \approx 600{,}000 \rightarrow 2^{20}\text{ (next power of 2)}.
\end{equation}
With blowup 16: $M_{\mathrm{outer}} = 2^{24}$. The LDE trace alone is
$\sim 50 \times 2^{24} \times 8 = 6.7$~GB, far exceeding the L3 cache
of any commodity CPU.

\subsubsection{Flaw 5: Wrong composition order}

The pipeline runs tx proofs $\to$ aggregation (recursion) $\to$ block commitment.
The aggregation is redundant: the block commitment circuit already binds
the transaction set.

\subsection{Clock cycle budget from first principles}

\subsubsection{Field operation costs on \texttt{hegemon-prover}}

Goldilocks arithmetic on a modern Xeon with AVX2 (4-wide SIMD packing,
\texttt{p3-goldilocks} implementation):
\begin{center}
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Operation} & \textbf{Amortized cycles} & \textbf{Notes} \\
\midrule
Field multiply & 2.5 & Barrett reduction after \texttt{mulhi}/\texttt{mullo} \\
Field add/sub & 0.5 & Conditional subtraction of $p$ \\
NTT butterfly & 3.5 & 1 mul $+$ 2 add $+$ twiddle lookup \\
Poseidon2 permutation & 2{,}420 & $736 \times 2.5 + 1159 \times 0.5$ \\
\bottomrule
\end{tabular}
\end{center}

Machine-level budget for 32 cores at $f \in [2.5, 3.0]$~GHz:
\begin{equation}
C_{\mathrm{block}} = 32 \times f \times 6 \in
[4.80 \times 10^{11},\; 5.76 \times 10^{11}]
\;\text{cycles per 6-second block.}
\end{equation}

Per-transaction budget as a function of TPS $T$ (with $n = 6T$ tx/block):
\begin{equation}
C_{\mathrm{tx}}(T) = \frac{C_{\mathrm{block}}}{6T}
= \frac{32 f}{T} \;\text{cycles/tx.}
\end{equation}

\begin{center}
\begin{tabular}{@{}rrr@{}}
\toprule
\textbf{TPS} & \textbf{tx/block} & \textbf{Budget (cycles/tx)} \\
\midrule
0.1 & $\sim$1 & $\le 5.76 \times 10^{11}$ \\
1 & 6 & $9.6 \times 10^{10}$ \\
10 & 60 & $9.6 \times 10^{9}$ \\
100 & 600 & $9.6 \times 10^{8}$ \\
1{,}000 & 6{,}000 & $9.6 \times 10^{7}$ \\
\bottomrule
\end{tabular}
\end{center}

\subsubsection{Single transaction proof budget}

For domain $M = 2^{17}$ (trace $N = 2^{13}$, blowup $b = 16$), width $W = 132$:

\paragraph{LDE/FFT.}
Each column requires a coset-FFT from $N$ to $M$ points. Dominant term:
\begin{equation}
B_{\mathrm{NTT}} = W \times \frac{M}{2} \times \log_2 M
= 132 \times 65{,}536 \times 17
= 147.0 \times 10^6 \text{ butterflies}.
\end{equation}
At 3.5 cycles/butterfly: $514.4 \times 10^6$~cycles single-threaded.
On 32 cores: $\sim 16.1 \times 10^6$~cycles ($\sim 6.4$~ms at 2.5~GHz).

\paragraph{Merkle commitment.}
Three oracle trees (trace, quotient, coefficient) plus 13 FRI commit-phase
trees. Total internal nodes:
\begin{align}
N_{\mathrm{nodes}} &\approx 2(M-1) + (M-1) + \sum_{i=1}^{13}(M/2^i - 1)
\nonumber\\
&= 2 \times 131{,}071 + 131{,}071 + 131{,}043
= 524{,}256.
\end{align}
Each compression uses one Poseidon2 permutation (2{,}420 cycles amortized):
\begin{equation}
C_{\mathrm{Merkle}} = 524{,}256 \times 2{,}420 = 1.269 \times 10^9
\text{ cycles single-threaded.}
\end{equation}
On 32 cores: $\sim 39.6 \times 10^6$~cycles ($\sim 15.9$~ms).

Note: Merkle tree construction is \emph{not} embarrassingly parallel.
At depth $d$ of the tree, there are only $M / 2^{d+1}$ independent nodes.
The top $\log_2(32) = 5$ levels are sequential, adding $\sim 5 \times
2{,}420 = 12{,}100$~cycles of serial latency. In practice, Plonky3's
parallel Merkle builder achieves $\sim 60\%$ of ideal scaling, so the
realistic estimate is $\sim 66 \times 10^6$~cycles ($\sim 26$~ms).

\paragraph{Constraint evaluation and quotient computation.}
Evaluate the AIR constraints over $M = 131{,}072$ evaluation points,
then divide by the zerofier to obtain quotient columns. The transaction AIR
has $\sim 150$ constraints of degree $\le 8$, each requiring $\sim 15$ field
operations on average (multiplied-out polynomial evaluations):
\begin{equation}
C_{\mathrm{constraints}} \approx 131{,}072 \times 150 \times 15 \times 2.5
= 737.3 \times 10^6 \text{ cycles single-threaded.}
\end{equation}
On 32 cores: $\sim 23 \times 10^6$~cycles ($\sim 9.2$~ms).

\paragraph{FRI query phase.}
32 queries, each opening 3 Merkle paths of depth 17 plus FRI fold
evaluations across 13 rounds:
\begin{equation}
C_{\mathrm{FRI}} \approx 32 \times 3 \times 17 \times 2{,}420
+ 32 \times 13 \times 100
= 3.96 \times 10^6 + 4.16 \times 10^4
\approx 4.0 \times 10^6 \text{ cycles.}
\end{equation}
This is negligible relative to other phases.

\paragraph{Total single-tx proof budget.}
\begin{equation}\label{eq:single-tx-budget}
C_{\mathrm{tx,prove}} \approx
\underbrace{16.1\text{M}}_{\text{LDE}} +
\underbrace{66\text{M}}_{\text{Merkle}} +
\underbrace{23\text{M}}_{\text{constraints}} +
\underbrace{4\text{M}}_{\text{FRI}}
= 109 \times 10^6 \;\text{cycles (32 cores).}
\end{equation}
Wall-clock at 2.5~GHz: $\sim 44$~ms per single-tx proof. With real-world
overheads (memory latency, OS scheduling, Rayon task overhead):
\begin{equation}
\boxed{T_{\mathrm{tx,single}} \approx 100\text{--}300\;\text{ms on 32 cores.}}
\end{equation}

\subsubsection{Batch-16 proof budget}

For $k=16$ transactions, $N_{\mathrm{batch}} = 2^{17} = 131{,}072$ rows.
After blowup $b = 16$: $M = 2^{21} = 2{,}097{,}152$.

\paragraph{LDE/FFT.}
\begin{equation}
B_{\mathrm{NTT}}^{(16)} = 132 \times \frac{2^{21}}{2} \times 21
= 132 \times 1{,}048{,}576 \times 21
= 2.904 \times 10^9 \text{ butterflies.}
\end{equation}
At 3.5 cycles/butterfly: $10.16 \times 10^9$~cycles.
On 32 cores: $317 \times 10^6$~cycles ($\sim 127$~ms at 2.5~GHz).

\paragraph{Memory bandwidth constraint.}
LDE trace size: $132 \times 2^{21} \times 8 = 2.21$~GB. A typical dual-socket
Xeon with DDR5 delivers $\sim 150$--$200$~GB/s sustained bandwidth. The NTT
requires $\sim 3$--$5$ passes over this data (butterfly stages with non-unit
stride). Total data movement:
\begin{equation}
D_{\mathrm{NTT}} \approx 4 \times 2.21\;\text{GB} = 8.84\;\text{GB}.
\end{equation}
At 175~GB/s: $\sim 50$~ms for data movement alone. The NTT is
\textbf{compute-bound, not bandwidth-bound} at this domain size, because the
arithmetic intensity ($\sim 3.5$ ops per 8-byte read) exceeds the machine's
compute-to-bandwidth ratio.

\paragraph{Merkle commitment.}
At $M = 2^{21}$: total internal nodes $\approx 6.3 \times 10^6$ across all
trees (trace $+$ quotient $+$ FRI phases with $c = 17$ rounds).
\begin{equation}
C_{\mathrm{Merkle}}^{(16)} = 6.3 \times 10^6 \times 2{,}420
= 15.2 \times 10^9 \;\text{cycles.}
\end{equation}
On 32 cores at 60\% efficiency: $\sim 793 \times 10^6$~cycles ($\sim 317$~ms).

For Merkle trees, the bottleneck shifts: each leaf is $132 \times 8 = 1{,}056$
bytes of trace data that must be hashed into a 48-byte digest. The
\emph{hashing} of leaves (not internal compression) dominates:
\begin{equation}
\text{Leaf hashing:}\;\;
2{,}097{,}152 \;\text{leaves} \times
\lceil 132 / 6 \rceil = 22 \;\text{absorb cycles each}
= 46.1 \times 10^6 \;\text{Poseidon2 perms.}
\end{equation}
At 2{,}420 cycles/perm on 32 cores at 60\%: $\sim 5.8 \times 10^9$~cycles
($\sim 2.3$~s). \textbf{This is the dominant cost.}

However, this analysis reveals an optimization: the trace commitment in
Plonky3 uses the \texttt{MerkleTreeMmcs} which hashes column \emph{packs},
not individual elements. With column packing and vectorized Poseidon2,
the practical cost for the full Merkle commit phase is closer to
$400$--$600$~ms on 32 cores.

\paragraph{Total batch-16 proof budget.}
\begin{equation}\label{eq:batch16-budget}
\boxed{T_{\mathrm{batch16}} \approx 1.5\text{--}3.0\;\text{s on 32 cores.}}
\end{equation}

The range reflects uncertainty in Merkle tree implementation efficiency.
With optimized leaf packing (which \texttt{p3-merkle-tree} implements):
closer to 1.5~s. With na\"ive per-element hashing: up to 3.0~s.

\subsubsection{Recursion overhead: why it costs $>50\times$ more}

For the recursion circuit verifying $n = 6$ inner proofs, the outer circuit
must encode $n$ copies of the FRI verifier. Each verifier copy adds
$\sim 100{,}000$ constraint-rows to the outer trace (accounting for
the 64{,}192 S-box evaluations, challenge derivation, and Merkle
path reconstruction). The outer trace has:
\begin{align}
N_{\mathrm{outer}} &\approx 6 \times 100{,}000 = 600{,}000
\rightarrow 2^{20}, \\
M_{\mathrm{outer}} &= 2^{20} \times 16 = 2^{24} = 16{,}777{,}216, \\
\text{LDE trace:}\;\; &\approx 50 \times 2^{24} \times 8 = 6.71\;\text{GB.}
\end{align}

The NTT over this domain:
\begin{equation}
B_{\mathrm{NTT}}^{\mathrm{outer}} = 50 \times 2^{23} \times 24
= 10.07 \times 10^9 \;\text{butterflies}
= 35.2 \times 10^9 \;\text{cycles.}
\end{equation}
On 32 cores: $1.1 \times 10^9$~cycles ($\sim 440$~ms). But memory bandwidth
now matters: 6.71~GB $\times$ 4 passes $= 26.8$~GB at 175~GB/s $= 153$~ms.
The operations are pipelined, so wall-clock is $\max(440, 153) \approx 440$~ms
for NTT alone.

Merkle commitment is worse: $\sim 3 \times 2^{24} \times 2{,}420 = 122 \times
10^9$~cycles on 32 cores at 60\%: $\sim 6.3 \times 10^9$~cycles ($\sim
2.5$~s). Plus leaf hashing dominates again.

The total outer proof takes $\sim 5$--$15$~s on 32 cores, which exceeds the
6-second block time. In addition, the circuit \emph{build} (symbolic constraint
graph construction via \texttt{CircuitBuilder}) takes 100--465~s on a cold
miss, making the system non-viable.

\paragraph{The recursion tax, precisely.}
For 6 transactions, the choices are:
\begin{center}
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Approach} & \textbf{Wall-clock} & \textbf{Cycles (32 cores)} \\
\midrule
6 individual STARK proofs (parallel) & $\sim 300$~ms & $750 \times 10^6$ \\
1 batch-6 proof ($2^{16}$ domain) & $\sim 800$~ms & $2.0 \times 10^9$ \\
Recursion over 6 tx proofs (hot) & $\sim 5$--$15$~s & $12$--$38 \times 10^9$ \\
Recursion over 6 tx proofs (cold) & $\sim 100$--$465$~s & $250$--$1{,}160 \times 10^9$ \\
\bottomrule
\end{tabular}
\end{center}

Recursion is $17\times$--$50\times$ more expensive than batch proving for the
same transaction set, and $1{,}500\times$ more expensive on cold cache misses.

\subsection{Seamless scaling from 0.1 to $>$1000 TPS}

The goal is a single architecture that scales from genesis-block emptiness to
world-commerce throughput without mode switches, configuration changes, or
hardware replacement. The key insight: \textbf{the architecture must transition
smoothly through throughput regimes, not jump between them.}

\subsubsection{Regime 0: Bootstrap ($\le 0.1$ TPS, $n \le 1$ tx/block)}

\paragraph{Workload:} At most 1 transaction per 6-second block. The prover is
mostly idle.

\paragraph{Proving strategy:} Single transaction STARK proof (domain $2^{17}$,
width 132). No batching needed---the overhead of padding a batch of 1 to $2^{13}$
rows is identical to a single-tx proof.

\paragraph{Cycle budget:}
\begin{equation}
T_{\mathrm{prove}} \approx 200\;\text{ms} \ll 6\;\text{s.}
\end{equation}
Utilization: $\sim 3\%$. The remaining $\sim 5.8$~s is pure slack.

\paragraph{Use idle time:} Prebuild preprocessed traces for batch sizes
$k \in \{1, 2, 4, 8, 16\}$ (5 configurations). Each preprocessed trace
takes $<1$~s to generate. This eliminates all cold paths for future regimes.

\subsubsection{Regime 1: Early commerce (1 TPS, $n = 6$ tx/block)}

\paragraph{Proving strategy:} One batch-6 proof. Trace: $6 \times 8192 =
49{,}152 \rightarrow 2^{16} = 65{,}536$ rows. Domain: $2^{16} \times 16 =
2^{20}$.

\paragraph{Cycle budget:}
\begin{align}
T_{\mathrm{batch}} &\approx 800\;\text{ms} \quad(2^{20}\text{ domain, 32 cores}), \\
T_{\mathrm{commit}} &\approx 200\;\text{ms} \quad(\text{block commitment, }2^{11}\text{ domain}), \\
T_{\mathrm{serialize}} &\approx 100\;\text{ms}, \\
T_{\mathrm{total}} &\approx 1.1\;\text{s} \ll 6\;\text{s.}
\end{align}
Utilization: $\sim 18\%$. Comfortable margin for variance, OS jitter,
and network delays.

\paragraph{Verification:} One batch STARK proof $+$ one commitment proof.
Verification time: $\sim 15$~ms total. Negligible vs.\ 6-second block time.

\paragraph{Proof payload:} One batch proof $\sim 100$~KB $+$ one commitment
proof $\sim 20$~KB $\approx 120$~KB per block. Acceptable for peer gossip.

\subsubsection{Regime 2: Production (10 TPS, $n = 60$ tx/block)}

\paragraph{Proving strategy:} Partition 60 transactions into $\lceil 60/16
\rceil = 4$ batches of $\le 16$. Generate 4 batch proofs in parallel, each
on 8 cores.

\paragraph{Cycle budget per batch (8 cores, $2^{17}$ domain):}
Scaling from 32 cores to 8 cores: NTT and constraint evaluation scale
linearly, but Merkle construction scales sub-linearly (tree parallelism
is limited at lower levels). Practical estimate: $4\times$ slower than
32-core, i.e.\ $\sim 2\times$ the 32-core time due to Amdahl's law
for serial fractions:
\begin{equation}
T_{\mathrm{batch}}^{(8\text{ cores})} \approx 2.0 \times T_{\mathrm{batch}}^{(32\text{ cores})}
\approx 3.0\text{--}5.0\;\text{s.}
\end{equation}

This is tight against the 6-second block time. A better allocation:
run 2 batches in parallel on 16 cores each, then the other 2 batches:
\begin{align}
\text{Wave 1:} &\quad 2\;\text{batches} \times 16\;\text{cores}
= 32\;\text{cores}, \quad T \approx 2.0\;\text{s}, \\
\text{Wave 2:} &\quad 2\;\text{batches} \times 16\;\text{cores}
= 32\;\text{cores}, \quad T \approx 2.0\;\text{s}, \\
T_{\mathrm{commit}} &\approx 0.3\;\text{s}, \quad
T_{\mathrm{total}} \approx 4.3\;\text{s} \le 6\;\text{s.} \checkmark
\end{align}

\paragraph{Alternative: 2 batches of 30.}
If the batch circuit is extended to support $k = 32$ (trace $= 2^{18}$,
domain $= 2^{22}$), then 60 tx $\rightarrow$ 2 batches of 30.
Each batch on 16 cores: $\sim 2.5$~s. Total $\approx 2.8$~s.
\textbf{This is the recommended approach for 10 TPS.}

\paragraph{Proof payload:} 2--4 batch proofs $\approx 200$--800~KB $+$
one commitment proof $\sim 25$~KB. Total $\sim 400$~KB.

\paragraph{Verification:} 2--4 batch verifications + 1 commitment verify
$\approx 25$--40~ms. Invisible at consensus timescale.

\subsubsection{Regime 3: High throughput (100 TPS, $n = 600$ tx/block)}

\paragraph{Why batch-only breaks down.}
With $k = 16$: $\lceil 600/16 \rceil = 38$ batch proofs. With 32 cores
constrained to one machine, batches must be serialized in waves:
$\lceil 38 / 2 \rceil = 19$ waves of 2 batches on 16 cores each,
total $\approx 19 \times 2 = 38$~s $\gg 6$~s.

Even with $k = 64$ (requiring batch AIR extension): $\lceil 600/64 \rceil =
10$ batch proofs with domain $2^{22}$. Each batch on 32 cores: $\sim 5$~s.
Sequential: $50$~s. Need $\ge 9$ parallel provers to fit in 6~s.

\paragraph{This is where recursion becomes justified.}
The key difference from the current broken recursion: the inner proofs
are \textbf{batch proofs} of identical shape, not variable-count
transaction proofs.

\paragraph{Architecture for 100 TPS:}
\begin{enumerate}[leftmargin=2em]
  \item \textbf{Batch layer:} 38 batch-16 proofs, generated on a
        prover cluster or GPU (see Section~\ref{sec:hardware}).
  \item \textbf{Merge layer:} Fixed-arity merge circuit (arity $a = 8$)
        that verifies 8 batch proofs of identical shape. This produces
        $\lceil 38/8 \rceil = 5$ merge proofs.
  \item \textbf{Root layer:} One merge proof over the 5 merge proofs
        (pad to 8 with trivial proofs).
  \item \textbf{Commitment:} Block commitment proof over all 600
        statement hashes.
\end{enumerate}

\paragraph{Why this recursion works when the current one does not:}
\begin{itemize}[leftmargin=2em]
  \item All inner proofs have identical shape: same domain $2^{21}$, same
        width 132, same FRI parameters. One circuit shape, cached once.
  \item The merge circuit verifies 8 batch proofs, each $\sim 50{,}000$
        constraint rows $\Rightarrow$ outer trace $\sim 400{,}000
        \rightarrow 2^{19}$ rows, domain $2^{23}$,
        LDE $\sim 50 \times 2^{23} \times 8 = 3.4$~GB.
  \item Merge proof on 32 cores: $\sim 4$--$6$~s (compute) or
        $\sim 0.5$--$1$~s (GPU).
  \item Tree depth: $\lceil \log_8 38 \rceil = 2$ levels. Total recursion
        latency: $2 \times T_{\mathrm{merge}}$ (pipelined, each level
        waits for the previous).
\end{itemize}

\paragraph{Timing for 100 TPS on a prover cluster (4$\times$ 32-core machines):}
\begin{center}
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Phase} & \textbf{Wall-clock} & \textbf{Hardware} \\
\midrule
Batch proving (38 $\div$ 4 machines) & $\sim 3.0$~s & 4 machines, 10 batches each \\
Merge level 1 (5 merge proofs) & $\sim 5.0$~s & 2 machines \\
Merge level 2 (1 root proof) & $\sim 5.0$~s & 1 machine \\
Commitment proof & $\sim 0.5$~s & 1 machine \\
\midrule
Critical path (pipelined) & $\sim 5.5$~s & \checkmark\ fits in 6~s \\
\bottomrule
\end{tabular}
\end{center}

Without pipelining: $13.5$~s $\gg 6$~s.
\textbf{Pipelining is essential}: start merge level 1 as soon as the first
8 batch proofs are ready, not after all 38 complete.

\paragraph{Timing for 100 TPS with GPU proving (see Section~\ref{sec:hardware}):}
\begin{center}
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Phase} & \textbf{Wall-clock} & \textbf{Hardware} \\
\midrule
Batch proving (38 proofs) & $\sim 1.5$~s & 2 GPUs, 19 batches each \\
Merge level 1 (5 proofs) & $\sim 1.0$~s & 1 GPU \\
Merge level 2 (1 proof) & $\sim 1.0$~s & 1 GPU \\
Commitment proof & $\sim 0.2$~s & CPU \\
\midrule
Critical path (pipelined) & $\sim 2.7$~s & ample margin \\
\bottomrule
\end{tabular}
\end{center}

\subsubsection{Regime 4: World commerce ($>$1000 TPS, $n > 6{,}000$ tx/block)}

At 1000 TPS: 6{,}000 transactions per block $\rightarrow$ 375 batch-16 proofs.

\paragraph{Recursion tree:}
\begin{align}
\text{Level 0 (leaves):} &\quad 375\;\text{batch proofs}, \\
\text{Level 1 (merge-8):} &\quad \lceil 375/8 \rceil = 47\;\text{merge proofs}, \\
\text{Level 2 (merge-8):} &\quad \lceil 47/8 \rceil = 6\;\text{merge proofs}, \\
\text{Level 3 (root):} &\quad 1\;\text{root proof.}
\end{align}
Tree depth: 3. Each level uses the same fixed merge circuit.

\paragraph{GPU cluster sizing (H100-class):}
\begin{itemize}[leftmargin=2em]
  \item Per batch-16 proof on H100: $\sim 80$~ms (NTT $+$ Merkle $+$ FRI,
        leveraging 3.35~TB/s HBM3 bandwidth and 16{,}896 CUDA cores).
  \item Per merge-8 proof on H100: $\sim 400$~ms (outer domain $2^{23}$,
        large trace).
  \item Batch layer: $375 \div 8\;\text{GPUs} = 47$ batches/GPU $\times$
        80~ms $= 3.76$~s.
  \item Merge level 1: $47 \div 4\;\text{GPUs} = 12$ merges/GPU $\times$
        400~ms $= 4.8$~s.
  \item Merge level 2: $6 \times 400$~ms $= 2.4$~s (1 GPU).
  \item Root: 400~ms (1 GPU).
\end{itemize}

\paragraph{Pipeline critical path:} With full pipelining (start next level's
proofs as soon as arity-count inputs are ready):
\begin{equation}
T_{\mathrm{critical}} \approx T_{\mathrm{first~8~batches}}
+ 2 \times T_{\mathrm{merge}} + T_{\mathrm{root}}
= 640\;\text{ms} + 800\;\text{ms} + 400\;\text{ms}
= 1.84\;\text{s} \ll 6\;\text{s.}
\end{equation}

\paragraph{Hardware cost:} 8 H100 GPUs ($\sim$\$250K)
for the batch layer, 4 for merge layer 1, 1 for merge level 2 and root.
Total: $\sim$13 GPUs. At H100 pricing ($\sim$\$30K each): $\sim$\$400K
capital for a system handling \$100B+/day in transaction volume (comparable
to SWIFT). This is economically rational.

\paragraph{At 10{,}000 TPS:} 60{,}000 tx/block $\rightarrow$ 3{,}750
batch proofs. Tree depth 4. Batch layer: $3{,}750 \div 64\;\text{GPUs}
\approx 59$ batches/GPU $\times$ 80~ms $= 4.7$~s. Still fits in 6~s
with $\sim$80 GPUs total. Linear hardware scaling.

\subsection{The rational architecture}

\subsubsection{Design principle: phase transitions, not mode switches}

The correct system does not have ``modes.'' It has a single pipeline
with parameters that adapt to load:

\begin{equation}
\text{Pipeline} = \text{Batch}(k) \;\|\; \text{Merge}^{d}(a)
\;\|\; \text{Commit}(n),
\end{equation}
where $k$ is batch size (16--64), $a$ is merge arity (4--8), and $d$ is
recursion depth ($0$ for $\le$10 TPS, $1$--$2$ for 100 TPS, $3$--$4$ for 1000+ TPS).

At low load ($d = 0$): no recursion at all. The pipeline is just
$\text{Batch}(k) \;\|\; \text{Commit}(n)$.

At high load: recursion engages automatically when the number of batch
proofs exceeds a threshold (e.g., $>8$).

\subsubsection{The batch-first pipeline}

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Partition:} $n$ transactions $\to$ $\lceil n/k \rceil$
        batches of $\le k$ transactions.

  \item \textbf{Batch prove:} Generate each batch STARK proof independently.
        Fully parallel. No inter-batch dependencies. Fixed AIR shape.
        Preprocessed trace computed once at startup.

  \item \textbf{Merge (conditional):} If batch count $> a$, build a
        recursion tree. Fixed-arity merge circuit verifies $a$ identical-shape
        batch proofs. Each merge level uses the same circuit.
        One cached circuit shape, zero cold misses.

  \item \textbf{Commit:} Run \texttt{CommitmentBlockProverP3} over all
        $n$ statement hashes.

  \item \textbf{Assemble:} Block payload contains either:
  \begin{itemize}[leftmargin=2em]
    \item (low load): $\lceil n/k \rceil$ batch proofs $+$ commitment proof.
    \item (high load): 1 root recursive proof $+$ commitment proof $+$ metadata.
  \end{itemize}
\end{enumerate}

\subsubsection{Soundness analysis}

For $B = \lceil n/k \rceil$ batch proofs, each with FRI soundness
$\varepsilon_1 = 2^{-128}$:
\begin{equation}
\varepsilon_{\mathrm{block}} \le B \cdot \varepsilon_1 + \varepsilon_{\mathrm{commit}}.
\end{equation}
For $B = 38$ (100 TPS): $\varepsilon_{\mathrm{block}} \le 38 \cdot 2^{-128}
= 2^{-122.8}$. Still well within the 128-bit security class
($< 2^{-120}$).

For $B = 375$ (1000 TPS): $\varepsilon_{\mathrm{block}} \le 375 \cdot 2^{-128}
= 2^{-119.4}$. Marginal. If this is unacceptable, increase FRI queries from
32 to 36, giving $\varepsilon_{\mathrm{FRI}} = 2^{-144}$ per proof, and
$375 \cdot 2^{-144} = 2^{-135.4}$.

With recursion (tree depth $d$, arity $a$): each level adds one union
bound. Total:
\begin{equation}
\varepsilon_{\mathrm{recursive}} \le B \cdot \varepsilon_1
+ d \cdot \varepsilon_{\mathrm{merge}} + \varepsilon_{\mathrm{commit}}.
\end{equation}
For $d = 3$, $\varepsilon_{\mathrm{merge}} = 2^{-128}$:
$\varepsilon_{\mathrm{recursive}} \le 2^{-119.4} + 3 \cdot 2^{-128}
\approx 2^{-119.4}$.

\subsection{Ideal hardware for STARK recursion}\label{sec:hardware}

\subsubsection{Why 32-core CPU is poor for recursion but good for batch proving}

The 32-core \texttt{hegemon-prover} (256~GB DDR5) has:
\begin{itemize}[leftmargin=2em]
  \item Peak bandwidth: $\sim 200$~GB/s (8-channel DDR5-4800).
  \item L3 cache: $\sim 48$--$105$~MB (depending on Xeon generation).
  \item Peak FLOPs: $32 \times 2 \times f \times 8$ (AVX2, 4-wide Goldilocks)
        $\approx 2.6$~T~ops/s at 2.5~GHz.
\end{itemize}

For a batch-16 proof ($M = 2^{21}$, trace $= 2.21$~GB): the trace
\textbf{spills the L3 cache} but the access pattern is streaming
(sequential column reads during NTT), which the hardware prefetcher
handles well. The CPU is efficient here.

For a recursion merge proof ($M = 2^{23}$--$2^{24}$, trace $= 3.4$--$6.7$~GB):
the trace is $30$--$130\times$ larger than L3. The Merkle tree construction
requires \textbf{random-access reads} (parent $\to$ children scatter/gather),
which thrash the cache hierarchy. CPU efficiency drops to $\sim 20$--$30\%$
of peak.

\paragraph{The diagnosis:} CPUs are compute-efficient for small-to-medium
proofs ($M \le 2^{21}$) where the working set fits in or streams through
the cache hierarchy. Recursion proofs have $M \ge 2^{23}$, where
\textbf{memory bandwidth and cache capacity} become the binding constraints.

\subsubsection{GPU: the natural recursion substrate}

Modern HPC GPUs (NVIDIA H100, AMD MI300X) are designed for exactly the
workload profile that recursion demands:

\begin{center}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Metric} & \textbf{Xeon 32-core} & \textbf{H100 SXM} &
  \textbf{Ratio} \\
\midrule
Memory bandwidth & 200~GB/s & 3{,}350~GB/s & $16.8\times$ \\
Memory capacity & 256~GB DDR5 & 80~GB HBM3 & $0.31\times$ \\
Compute (Goldilocks mul) & $\sim 2.6$~T ops/s & $\sim 30$~T ops/s & $11.5\times$ \\
L2/cache & 105~MB & 50~MB & $0.48\times$ \\
NTT parallelism & 32 threads & 16{,}896 CUDA cores & $528\times$ \\
\bottomrule
\end{tabular}
\end{center}

The GPU's advantages are precisely where the CPU fails:
\begin{enumerate}[leftmargin=2em]
  \item \textbf{NTT:} The $M \log M$ butterfly pattern is embarrassingly
        parallel. GPU NTT implementations \cite{ingonyama2024} achieve
        $10$--$20\times$ speedup over 32-core CPU for $M \ge 2^{22}$.
  \item \textbf{Merkle tree:} Each tree level's hashing is independent.
        With 16{,}896 cores, even the top levels of a $2^{24}$-leaf tree
        have enough parallelism.
  \item \textbf{Poseidon2:} The permutation is $736$ multiplications $+$
        $1{,}159$ additions per invocation, with no data-dependent branching.
        This maps perfectly to GPU SIMT execution.
\end{enumerate}

The GPU's disadvantage---smaller total memory (80~GB vs.\ 256~GB)---is
irrelevant for merge proofs: even the largest recursion trace ($2^{24}$
domain, 50 columns) is $\sim 6.7$~GB, well within HBM capacity.

\paragraph{Expected GPU proving times:}
\begin{center}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Proof type} & \textbf{32-core CPU} & \textbf{H100 GPU} \\
\midrule
Single tx ($M = 2^{17}$) & 200~ms & 30~ms \\
Batch-16 ($M = 2^{21}$) & 2.0~s & 150~ms \\
Merge-8 ($M = 2^{23}$) & 5.0~s & 500~ms \\
Recursion root ($M = 2^{24}$) & 10.0~s & 800~ms \\
\bottomrule
\end{tabular}
\end{center}

These estimates are consistent with published GPU STARK prover benchmarks:
Icicle/Ingonyama reports $\sim 8\times$ speedup for Goldilocks NTT on A100
\cite{ingonyama2024}, and the Stwo (Circle STARK) GPU prover achieves
sub-second proofs for $2^{22}$-row traces \cite{stwo2024}.

\subsubsection{Heterogeneous pipeline: CPU batch $+$ GPU recursion}

The optimal hardware architecture for Hegemon at scale:

\begin{enumerate}[leftmargin=2em]
  \item \textbf{CPU prover (existing 32-core machine):} Handles batch
        proving for $\le 10$ TPS. No recursion needed. The batch circuit
        runs on compile-time-fixed AIR shapes, and CPU is efficient for
        medium-domain proofs ($M \le 2^{21}$).

  \item \textbf{GPU coprocessor (H100 or equivalent):} Engaged when batch
        count exceeds the merge arity threshold. Handles merge proofs
        ($M = 2^{23}$--$2^{24}$) where bandwidth dominates.

  \item \textbf{Prover cluster (multiple CPU or GPU nodes):} Required only
        at $>100$ TPS, when the batch layer itself needs parallelization
        across machines.
\end{enumerate}

\paragraph{Scaling table by TPS regime:}
\begin{center}
\begin{tabular}{@{}rllrl@{}}
\toprule
\textbf{TPS} & \textbf{Batch HW} & \textbf{Merge HW} & \textbf{Time} &
  \textbf{Cost} \\
\midrule
0.1 & 1 CPU (32c) & none & $<1$~s & \$5K \\
1 & 1 CPU (32c) & none & $\sim 1$~s & \$5K \\
10 & 1 CPU (32c) & none & $\sim 4$~s & \$5K \\
100 & 4 CPU (32c) & 1 GPU & $\sim 5$~s & \$50K \\
1{,}000 & 8 GPU & 4 GPU & $\sim 2$~s & \$400K \\
10{,}000 & 64 GPU & 16 GPU & $\sim 5$~s & \$2.5M \\
\bottomrule
\end{tabular}
\end{center}

Hardware cost scales sub-linearly with TPS because batch proving is
$O(n)$ total work but $O(1)$ depth with enough parallelism, and recursion
adds only $O(\log n)$ depth.

\subsection{Extended quantitative comparison}

\begin{table}[h]
\centering
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Metric} & \textbf{Current} & \textbf{Batch-only} &
  \textbf{Batch$+$merge} \\
& \textbf{(recursion)} & \textbf{($\le$10 TPS)} &
  \textbf{($>$10 TPS)} \\
\midrule
Circuit build (cold) & 100--465~s & 0~s & $<1$~s (one merge shape) \\
Hot time (6 tx) & $>$5~s & 1.1~s & --- \\
Hot time (60 tx) & infeasible & 4.3~s & --- \\
Hot time (600 tx) & infeasible & infeasible & 5.5~s (4 CPUs) \\
Cache key cardinality & $O(n_{\max})$ & $O(1)$ & $O(1)$ \\
Peak memory & $>$64~GB & $\sim$4~GB & $\sim$8~GB \\
Proofs per block & 1 & $\lceil n/16\rceil+1$ & 1 root $+$ 1 commit \\
Verifier cost & $<10$~ms & $<50$~ms & $<15$~ms \\
Proof payload & $\sim$200~KB & $\sim$400~KB & $\sim$250~KB \\
Soundness & $2^{-120}$ & $2^{-124}$ & $2^{-119}$ \\
Max TPS (1 machine) & $<0.2$ & $\sim$15 & $\sim$100 (w/ GPU) \\
\bottomrule
\end{tabular}
\caption{Full comparison across architectures.}
\end{table}

\subsection{What must change in the codebase}

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Immediate (Phase 1, ship now):} Wire
        \texttt{BatchTransactionProverP3::prove\_batch()} into the block
        production pipeline. Replace the aggregation circuit call. Use
        \texttt{CommitmentBlockProverP3} for block-level binding. Target:
        0.1--10 TPS on existing hardware.

  \item \textbf{Phase 2 (engineering, 2--4 weeks):} Extend batch size from
        16 to 32 or 64 transactions. Build a fixed-arity merge circuit
        (arity 8) that verifies 8 identical batch proofs. Prebuild the
        single merge circuit shape at node startup ($<1$~s). Target:
        10--100 TPS.

  \item \textbf{Phase 3 (GPU integration, 1--3 months):} Port the batch
        and merge provers to GPU using Icicle \cite{ingonyama2024} or
        a custom CUDA Goldilocks NTT$+$Poseidon2 kernel. The AIR constraints
        and FRI protocol are unchanged; only the prover backend is
        replaced. Target: 100--1000+ TPS.

  \item \textbf{Phase 4 (research, 6+ months):} Evaluate folding schemes
        (Nova \cite{nova2022}, HyperNova \cite{hypernova2023}) for the
        merge step. Folding avoids the STARK-in-STARK overhead by
        accumulating instances incrementally. Only justified if the merge
        circuit becomes the bottleneck above $\sim$1000 TPS.
\end{enumerate}

\subsection{On the prior document's tree-recursive proposal}

Section~9's tree-DAG recursion with fixed shapes is directionally correct for
$>$100 TPS but makes a critical error: \emph{it recurses over individual
transaction proofs, not batch proofs.} One batch-16 proof contains the same
validity information as 16 individual transaction proofs, but the merge
circuit verifying one batch proof processes it as a single proof shape rather
than 16 separate verifier gadgets. This is $16\times$ cheaper per transaction
at the merge level.

The proposal to ``promote the cache from thread-local to process-global'' is
treating a symptom of the wrong architecture rather than the root cause.
With a batch-first pipeline, there is no dynamic cache because all circuit
shapes are fixed at compile time.

\subsection{Summary: the path from genesis to world commerce}

The current codebase contains three correct primitives---the transaction AIR,
the batch circuit, and the block commitment circuit---composed in the wrong
order with a recursion layer that destroys performance. The fix is
architectural, not optimizational:

\begin{enumerate}[leftmargin=2em]
  \item \textbf{0.1--10 TPS:} Batch proving (existing code, unused). No
        recursion. CPU only. Budget: $<4.3$~s out of 6~s.
  \item \textbf{10--100 TPS:} Batch proving $+$ fixed-arity merge
        recursion. CPU batch $+$ GPU merge. Budget: $<5.5$~s.
  \item \textbf{100--1000 TPS:} GPU batch $+$ GPU recursive tree.
        Prover cluster. Budget: $<3$~s.
  \item \textbf{1000+ TPS:} GPU cluster scaling. Tree depth grows as
        $\log_8 n$. Hardware cost scales sub-linearly.
\end{enumerate}

Security is preserved at every stage: the same Poseidon2/Goldilocks/FRI
parameters, the same AIR constraints, the same 128-bit PQ soundness target.
The only variable is \emph{how proofs are composed}, not \emph{what they prove}.

\section*{References}
\begin{thebibliography}{99}

\bibitem{stark2018}
E. Ben-Sasson, I. Bentov, Y. Horesh, M. Riabzev.
\newblock \emph{Scalable, transparent, and post-quantum secure computational integrity}.
\newblock IACR ePrint 2018/046.
\newblock \url{https://eprint.iacr.org/2018/046}

\bibitem{deepfri2019}
E. Ben-Sasson, L. Goldberg, S. Kopparty, S. Saraf.
\newblock \emph{DEEP-FRI: Sampling outside the box improves soundness}.
\newblock arXiv:1903.12243.
\newblock \url{https://arxiv.org/abs/1903.12243}

\bibitem{fractal2019}
A. Chiesa, D. Ojha, N. Spooner.
\newblock \emph{Fractal: Post-Quantum and Transparent Recursive Proofs from Holography}.
\newblock IACR ePrint 2019/1076.
\newblock \url{https://eprint.iacr.org/2019/1076}

\bibitem{hypernova2023}
A. Kothapalli, S. Setty.
\newblock \emph{HyperNova: Recursive arguments for customizable constraint systems}.
\newblock IACR ePrint 2023/573.
\newblock \url{https://eprint.iacr.org/2023/573}

\bibitem{neo2025}
W. D. Nguyen, S. Setty.
\newblock \emph{Neo: Lattice-based folding scheme for CCS over small fields and pay-per-bit commitments}.
\newblock IACR ePrint 2025/294 (metadata index).
\newblock \url{https://dblp.org/rec/journals/iacr/NguyenS25}

\bibitem{superneo2026}
W. Nguyen, S. Setty.
\newblock \emph{Neo and SuperNeo: Post-quantum folding with pay-per-bit costs over small fields}.
\newblock IACR ePrint 2026/242 listing metadata.
\newblock \url{https://eprint-classic.github.io/2026.html}

\bibitem{poseidon22023}
L. Grassi, D. Khovratovich, M. Schofnegger.
\newblock \emph{Poseidon2: A Faster Version of the Poseidon Hash Function}.
\newblock IACR ePrint 2023/323.
\newblock \url{https://eprint.iacr.org/2023/323}

\bibitem{plonky3recbook}
Plonky3 Contributors.
\newblock \emph{The Plonky3 recursion book}.
\newblock \url{https://plonky3.github.io/Plonky3-recursion/introduction.html}

\bibitem{risc0repo}
RISC Zero.
\newblock \emph{risc0 repository}.
\newblock \url{https://github.com/risc0/risc0}

\bibitem{risc0docs}
RISC Zero.
\newblock \emph{risc0-zkvm crate documentation}.
\newblock \url{https://docs.rs/risc0-zkvm/}

\bibitem{nova2022}
A. Kothapalli, S. Setty, I. Tzialla.
\newblock \emph{Nova: Recursive Zero-Knowledge Arguments from Folding Schemes}.
\newblock CRYPTO 2022. IACR ePrint 2021/370.
\newblock \url{https://eprint.iacr.org/2021/370}

\bibitem{starkware2023}
StarkWare.
\newblock \emph{STARK Recursion in Practice: Recursive Proof Composition}.
\newblock StarkWare Blog, 2023.
\newblock \url{https://starkware.co/resource/recursive-starks/}

\bibitem{ingonyama2024}
Ingonyama.
\newblock \emph{ICICLE: GPU Acceleration for Zero-Knowledge Proofs}.
\newblock Ingonyama, 2024. Open-source CUDA library for NTT, MSM, and
hash-based commitments over Goldilocks and other fields.
\newblock \url{https://github.com/ingonyama-zk/icicle}

\bibitem{stwo2024}
StarkWare.
\newblock \emph{Stwo: Circle STARK Prover}.
\newblock StarkWare, 2024. High-performance STARK prover with GPU backend,
achieving sub-second proofs for large trace domains.
\newblock \url{https://github.com/starkware-libs/stwo}

\end{thebibliography}

\end{document}
